---
title: "week_6_lab"
author: "Lauren Skube"
date: "11/11/2020"
output: html_document
---

## Part 1: Attach packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

## Part 2: a rank-based test example (Mann Whitney U)
- create two sample vectors 
- use sample.int() to create random samples with integers 1 to x, of size =?, with replacement
- explore each vector with a histogram 
```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)

# Exploratory histogram 
hist(gp_1)
hist(gp_2)

```

### Why choose a rank-based test? 

1. Not clearly normally distributed from exploratory histograms 
2. Somewhat small sample size (n=15)
3. I've decided ranks (or medians) are a more valuable metric to compare these data 

## Perform Mann-Whitney U 
- to answer "is there a significant difference in ranks (medians) between gp_1 and gp_2? 
- using the wilcox.test() function 

```{r}
my_mwu <- wilcox.test(gp_1, gp_2) # Call in console to view 
```


### What does the p-value of 0.28 mean? 

- It means that if the null hypothesis is true (these samples were drawn from populations with the same median), there is a probability of 0.28 that we could have found median values *at least as different as ours* by chance.

## Part 3: Simple linear regression 
- use the penguins data set to explore the relationship between flipper length and body mass for penguins , including all 3 penguin species in the data set 

### Part 3(A): Look at data 
- make an exploratory scatterplot 

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```
### Questions to ask: 

1. does it look like a linear relationship makes sense? 
2. do we have any concerns about modeling as a linear relationship? 
3. any notable outliers? 
4. initial thoughts about homoscedasticity? 

### Part 3(B): Model it 
- using lm() function 

```{r}
# Linear model, stored as penguin_lm 
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

# Return the complete overview: 
summary(penguin_lm)
```
The multiple R2 value is 0.759- meaning that 75.9% of variance in body mass is explained by flipper length. 

### Part 3(C): Access model outputs

We can access the coefficients for the model using: 

- the slope is 49.69 (g/mm)
- the y-intercept is -5780.83
- the full equation is mass = 49.49*(flipperlength) + (-5780.83)


```{r}
# use the broom::tidy() function to get model outputs in nice data format 
penguin_lm_tidy <- broom::tidy(penguin_lm)

# Look at output by calling it in the console 

# Get the intercept: 
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int

# Get the flipper_length coefficient: 
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef

# Metrics at a glance: 
penguin_lm_out <- broom::glance(penguin_lm)
penguin_lm_out
  

```



